Automatic conversion of colloquial Finnish to standard Finnish

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
small cleanup
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58601 72bbbca6-d526-0410-a7d9-f06f51895060

\newcite{viinikka2013} describe the common meaning of the terms colloquial (\emph{puhekieli}) and standard (\emph{yleiskieli} or \emph{kirjakieli}) Finnish: standard language is regulated and is unified in morphology and vocabulary; colloquial language shows local and idiolectal variation, and has structures that are characteristic to spoken variety,  such as discourse particles and incomplete clauses.

Colloquial Finnish differs from standard Finnish in some key features:

Morphology shows dialectal variation, shortening and merging of morphemes, e.g. \emph{en min\"{a}} `not I' $\rightarrow$ \emph{emm\"{a}} (`I don't').

Syntax: due to the spontaneous nature of colloquial language, syntactic structures are common for spoken language. 
Prototypically, colloquial language is spoken and spontaneous, and standard is written, planned and structured. We illustrate the differences with the following example from our data set. Sentence 1 is the original colloquial version. The gloss shows the actual word-by-word translation, and the translation shows similar style and register in English.

@@

For the partial version, we did not change the syntactic structure, but changed only morphology and vocabulary. On the lexical level, the word \emph{seiska} `number 7' is colloquial style, and in the standard translation it is replaced by the ordinal \emph{seitsem\"{a}s} `seventh'.  Other changes in the partial version are common morphological or phonological, phenomena, such as reducing the diphthong in \emph{vain} $\rightarrow$ \emph{vaan} or shortening in  \emph{ett\"{a}} $\rightarrow$ \emph{et}. The original sentence and the partial translation are shown below, aligned word by word.

>>>>

\newcite{viinikka2013} describe the common meaning of the terms colloquial (\emph{puhekieli}) and standard (\emph{yleiskieli} or \emph{kirjakieli}) Finnish: standard language is unified in morphology and vocabulary, following the regulations of a language board; colloquial language shows local and idiolectal variation, and has structures that are characteristic to spoken variety,  such as discourse particles and incomplete clauses.

% Colloquial Finnish differs from standard Finnish in some key features:

% Morphology shows dialectal variation, shortening and merging of morphemes, e.g. \emph{en min\"{a}} `not I' $\rightarrow$ \emph{emm\"{a}} (`I don't').

% Syntax: due to the spontaneous nature of colloquial language, syntactic structures are common for spoken language. 
%Prototypically, colloquial language is spoken and spontaneous, and standard is written, planned and structured.

We illustrate the differences with the following example from our data set. Sentence 1 is the original colloquial version. The gloss shows the actual word-by-word translation, and the translation shows similar style and register in English.

@@

For the partial version, we changed only morphology and vocabulary. On the lexical level, the word \emph{seiska} `number 7' is colloquial style, and in the standard translation it is replaced by the ordinal \emph{seitsem\"{a}s} `seventh'.  Other changes in the partial version are common morphological or phonological, phenomena, such as reducing the diphthong in \emph{vain} $\rightarrow$ \emph{vaan} or shortening in  \emph{ett\"{a}} $\rightarrow$ \emph{et}. The original sentence and the partial translation are shown below, aligned word by word.
